---
fig.caption: yes
---

```{r setup, include=FALSE}
source("utilities/data.R")
source("utilities/plotLine.R")
source("utilities/CI.R")
source("utilities/modeling.R")
```

### Neural tube defects {#section-22}

At a $5\%$ level, there is `r ifelse(modeling_trend(data = anom_grp, value = "q00")[["pvalue"]] >= 0.05, "no significant", ifelse(modeling_trend(data = anom_grp, value = "q00")[["pvalue"]] < 0.05 & modeling_trend(data = anom_grp, value = "q00")[["coef"]] < 0, "a decreasing", "an increasing"))` trend with neural tube defects as a whole or within the sub-group encephalocele.

At a $5\%$ level, it was possible to see a slightly increasing and decreasing trend for anencephaly ($\hat{\beta_{1}} =$ `r scales::comma(modeling_trend(data = anom_ind, value = "q00")[["coef"]], accuracy = 0.01)`, $p-value=$ `r scales::comma(modeling_trend(data = anom_ind, value = "q00")$pvalue, accuracy = 0.01)`), and spina bifida ($\hat{\beta_{1}} =$ `r scales::comma(modeling_trend(data = anom_ind, value = "q05")$coef, accuracy = 0.01)`, $p-value=$ `r scales::comma(modeling_trend(data = anom_ind, value = "q05")$pvalue, accuracy = 0.01)`), respectively.

Overall, the prevalence rates have `r ifelse(modeling_trend(data = anom_grp[BrthYear > 2011], value = "q00")[["pvalue"]] >= 0.05, "been stable", ifelse(modeling_trend(data = anom_grp[BrthYear > 2011], value = "q00")[["pvalue"]] < 0.05 & modeling_trend(data = anom_grp[BrthYear > 2011], value = "q00")[["coef"]] < 0, "decreased", "increased"))` over the past 10 years - 2012-2021: $\hat{\beta_{1}} =$ `r scales::comma(modeling_trend(data = anom_grp[BrthYear > 2011], value = "q00")[["coef"]], accuracy = 0.01)`, p-value = `r scales::comma(modeling_trend(data = anom_grp[BrthYear > 2011], value = "q00")[["pvalue"]], accuracy = 0.01)`); $95\%$ CI:[`r scales::comma(modeling_trend(data = anom_grp[BrthYear > 2011], value = "q00")[["coef"]]-1.96*modeling_trend(data = anom_grp[BrthYear > 2011], value = "q00")[["std"]], accuracy = 0.01)`; `r scales::comma(modeling_trend(data = anom_grp[BrthYear > 2011], value = "q00")[["coef"]]+1.96*modeling_trend(data = anom_grp[BrthYear > 2011], value = "q00")[["std"]], accuracy = 0.01)`].

```{r nt-1, fig.cap= 'Neural tube defects - Total reported cases'}
dta <- anom_grp[grepl(tolower("q00"),
                   tolower(qcode)) ]
  
xy <- modeling_anal(anom_grp, "q00")

p1 <- plot_line(data = dta,
                var = "rate")

p1 %>% 
  add_trace(data = xy,
            x = ~BrthYear,
            y = ~rate,
            name = ifelse(var(dta$total_anom,
         na.rm = TRUE)/mean(dta$total_anom,
                            na.rm = TRUE) > 2,
         "Negative Binomial GLM fit",
         "Poisson GLM fit"),
            mode = "lines",
            line = list(color = "black")
            # marker = list(color = "black")
         )
```

```{r nt-2, fig.cap= 'Neural tube defects - Nova Scotia'}
# filter data
dta <- anom_grp_map[grepl(tolower("q00"),
                   tolower(qcode)) ]

# build map

```

```{r nt-3, fig.cap="Neural tube defects: Anencephaly, Encephalocele, Spina bifida - Total reported cases"}
## filtering neural tube defects
## count anomaly by year
## categorize them
# filtering total births: lvb + stillbirths

dta <- anom_ind[grepl(tolower("q00|q01|q05"),
                   tolower(qcode)) ]

plot_line(data = dta,
          var = "rate")
```