---
fig.caption: yes
params:
  year1: 2013
  year2: 2022
---

```{r setup, include=FALSE}
source("utilities/data.R")
source("utilities/plotLine.R")
source("utilities/CI.R")
source("utilities/modeling.R")
source("utilities/plotNSCDMap.R")

anom <- anom %>% 
  filter(!SrceIDs %in% 5, #remove MSI
         between(Birth_Year, params$year1, params$year2))
```

# Anomalies Overview {#chapter3}

## Neural tube defects {#section32}

At a $5\%$ level, there is `r ifelse(modeling_trend(data = anom, value = "gr101")[["pvalue"]] >= 0.05, "no significant", ifelse(modeling_trend(data = anom, value = "gr101")[["pvalue"]] < 0.05 & modeling_trend(data = anom, value = "gr101")[["coef"]] < 0, "a decreasing", "an increasing"))` trend for `r tolower(unique(subset(anom, tolower(Diag) %in% "gr101")[["cat"]]))` ($\hat{\beta_{1}} =$ `r scales::comma(modeling_trend(data = anom, value = "gr101")[["coef"]], accuracy = 0.001)`, p-value = `r ifelse(modeling_trend(data = anom, value = "gr101")[["pvalue"]] < 0.0001,"< 0.0001",scales::comma(modeling_trend(data = anom, value = "gr101")[["pvalue"]],accuracy = 0.001))`) for the period `r paste0(min(modeling_anal(data = anom, value = "gr101")[["Birth_Year"]])," - ",max(modeling_anal(data = anom, value = "gr101")[["Birth_Year"]]))`.

Over the past 11 years `r paste0("(",params$year1,"-",params$year1+10,")")`, `r ifelse(modeling_trend(data = anom, value = "gr101")[["pvalue"]] >= 0.05, "there is no evidence showing any trend", ifelse(modeling_trend(data = anom, value = "gr101")[["pvalue"]] < 0.05 & modeling_trend(data = anom, value = "gr101")[["coef"]] < 0, "there is evidence showing a decrease", "there is evidence showing an increase"))` in the prevalence rate of `r tolower(unique(anom %>% filter(tolower(Diag) %in% "gr101") %>% select(cat)))`. On average, the prevalence rate for each subsequent year is `r scales::comma(exp(modeling_trend(data = anom, value = "gr101")[["coef"]]), accuracy = 0.001)` times of the previous year's rate: (prevalence rate = exp(`r scales::comma(modeling_trend(data = anom, value = "gr101")[["coef"]], accuracy = 0.001)`) = `r scales::comma(exp(modeling_trend(data = anom, value = "gr101")[["coef"]]), accuracy = 0.001)` per 10,000 total births; $\hat{\beta_{1}} =$ `r scales::comma(modeling_trend(data = anom, value = "gr101")[["coef"]], accuracy = 0.001)`, p-value = `r ifelse(modeling_trend(data = anom, value = "gr101")[["pvalue"]] < 0.0001,"< 0.0001",scales::comma(modeling_trend(data = anom, value = "gr101")[["pvalue"]],accuracy = 0.001))`; $95\%$ CI for $\hat{\beta_{1}}$ and for the prevalence rate are given by $CI(95\%)_{\hat{\beta_{1}}}$:[`r scales::comma(modeling_trend(data = anom, value = "gr101")[["coef"]]-1.96*modeling_trend(data = anom, value = "gr101")[["std"]], accuracy = 0.001)`; `r scales::comma(modeling_trend(data = anom, value = "gr101")[["coef"]]+1.96*modeling_trend(data = anom, value = "gr101")[["std"]], accuracy = 0.001)`], and $CI(95\%)_{rate}$:[`r scales::comma(exp(modeling_trend(data = anom, value = "gr101")[["coef"]]-1.96*modeling_trend(data = anom, value = "gr101")[["std"]]), accuracy = 0.001)`; `r scales::comma(exp(modeling_trend(data = anom, value = "gr101")[["coef"]]+1.96*modeling_trend(data = anom, value = "gr101")[["std"]]), accuracy = 0.001)`], respectively).

```{r nt1, fig.cap=paste0("Neural tube defects - Prevalence - Nova Scotia - ", params$year1,"-",params$year2)}
dta <- subset(anom, tolower(Diag) %in% "gr101") %>% 
  select(CaseID, Birth_Year, count_brth_yr, Diag, cat) %>% 
  distinct() %>% 
  group_by(Birth_Year) %>% 
  mutate(total = n(),
         rate = total/count_brth_yr*10000) %>% 
  select(Birth_Year, rate, Diag, cat) %>% 
  distinct() %>% 
  ungroup()
  
mod1 <- modeling_anal(anom, "gr101")

if (knitr::is_html_output()){
p1 <- plot_line(data = dta,
                var = "rate")

p1 %>% 
  add_trace(data = mod1,
            x = ~Birth_Year,
            y = ~rate,
            name = ifelse(mod1[["model"]] %in% "nb",
                          "Negative Binomial GLM fit",
                          "Poisson GLM fit"),
            mode = "lines+markers",
            line = list(color = "black"),
            marker = list(color = "black"),
         hovertemplate = ~paste(
                    "<b>", cat, "-", Birth_Year, "</b>",
                    "<br> Estimated Prevalence:",
                    scales::comma(rate,
                                  accuracy = 0.001),
                    "<extra></extra>" # removes the trace name from the hover text
                  )
         )
} else if(knitr::is_latex_output()){
  mod1 <- mod1 %>% 
  mutate(cat = ifelse(model %in% "nb",
         "Negative Binomial GLM fit",
         ifelse(model %in% "pois",
         "Poisson GLM fit", cat))) %>% 
    select(-c(model,total,count_brth_yr))
  
  new_dta <- rbind(dta, mod1)
  
  plot_line_pdf(data = new_dta,
                var = "rate")
}
```

Based on the statistical analysis, at a $5\%$ level, it is possible to observe:

- `r unique(anom %>% filter(tolower(Diag) %in% "q00") %>% select(cat) %>% distinct() %>% pull())`: there `r ifelse(modeling_trend(data = anom, value = "q00")[["pvalue"]] >= 0.05, "is no significant", ifelse(modeling_trend(data = anom, value = "q00")[["pvalue"]] < 0.05 & modeling_trend(data = anom, value = "q00")[["coef"]] < 0, "is evidence of a decreasing", "is evidence of an increasing"))` trend ($\hat{\beta_{1}} =$ `r scales::comma(modeling_trend(data = anom, value = "q00")[["coef"]], accuracy = 0.001)`, p-value = `r ifelse(modeling_trend(data = anom, value = "q00")[["pvalue"]] < 0.0001,"< 0.0001",scales::comma(modeling_trend(data = anom, value = "q00")[["pvalue"]],accuracy = 0.001))`).

- `r unique(anom %>% filter(tolower(Diag) %in% "q01") %>% select(cat) %>% distinct() %>% pull())`: there `r ifelse(modeling_trend(data = anom, value = "q01")[["pvalue"]] >= 0.05, "is no significant", ifelse(modeling_trend(data = anom, value = "q01")[["pvalue"]] < 0.05 & modeling_trend(data = anom, value = "q01")[["coef"]] < 0, "is evidence of a decreasing", "is evidence of an increasing"))` trend ($\hat{\beta_{1}} =$ `r scales::comma(modeling_trend(data = anom, value = "q01")[["coef"]], accuracy = 0.001)`, p-value = `r ifelse(modeling_trend(data = anom, value = "q01")[["pvalue"]] < 0.0001,"< 0.0001",scales::comma(modeling_trend(data = anom, value = "q01")[["pvalue"]],accuracy = 0.001))`).

- `r unique(anom %>% filter(tolower(Diag) %in% "q05") %>% select(cat) %>% distinct() %>% pull())`: there `r ifelse(modeling_trend(data = anom, value = "q05")[["pvalue"]] >= 0.05, "is no significant", ifelse(modeling_trend(data = anom, value = "q05")[["pvalue"]] < 0.05 & modeling_trend(data = anom, value = "q05")[["coef"]] < 0, "is evidence of a decreasing", "is evidence of an increasing"))` trend ($\hat{\beta_{1}} =$ `r scales::comma(modeling_trend(data = anom, value = "q05")[["coef"]], accuracy = 0.001)`, p-value = `r ifelse(modeling_trend(data = anom, value = "q05")[["pvalue"]] < 0.0001,"< 0.0001",scales::comma(modeling_trend(data = anom, value = "q05")[["pvalue"]],accuracy = 0.001))`).

```{r nt11, fig.cap=paste0(unique(anom %>% filter(tolower(Diag) %in% "q00") %>% select(cat) %>% distinct() %>% pull()), " - Prevalence.")}

dta <- subset(anom, tolower(Diag) %in% c("q00")) %>% 
  select(CaseID, Birth_Year, count_brth_yr, Diag, cat) %>% 
  distinct() %>% 
  group_by(Birth_Year, Diag) %>% 
  mutate(total = n(),
         rate = total/count_brth_yr*10000) %>% 
  select(Birth_Year, rate, Diag, cat) %>% 
  distinct() %>% 
  ungroup()
  
mod1 <- modeling_anal(anom, "q00")

if (knitr::is_html_output()){
p1 <- plot_line(data = dta,
                var = "rate")

p1 %>% 
  add_trace(data = mod1,
            x = ~Birth_Year,
            y = ~rate,
            name = ifelse(mod1[["model"]] %in% "nb",
                          "Negative Binomial GLM fit",
                          "Poisson GLM fit"),
            mode = "lines+markers",
            line = list(color = "black"),
            marker = list(color = "black"),
         hovertemplate = ~paste(
                    "<b>", cat, "-", Birth_Year, "</b>",
                    "<br> Estimated Prevalence:",
                    scales::comma(rate,
                                  accuracy = 0.001),
                    "<extra></extra>" # removes the trace name from the hover text
                  )
         )
} else if(knitr::is_latex_output()){
  mod1 <- mod1 %>% 
  mutate(cat = ifelse(model %in% "nb",
         "Negative Binomial GLM fit",
         ifelse(model %in% "pois",
         "Poisson GLM fit", cat))) %>% 
    select(-c(model,total,count_brth_yr))
  
  new_dta <- rbind(dta, mod1)
  
plot_line_pdf(data = new_dta,
              var = "rate")
}
```

```{r nt12, fig.cap=paste0(unique(anom %>% filter(tolower(Diag) %in% "q01") %>% select(cat) %>% distinct() %>% pull())," - Prevalence.")}
dta <- subset(anom, tolower(Diag) %in% "q01") %>% 
  select(CaseID, Birth_Year, count_brth_yr, Diag, cat) %>% 
  distinct() %>% 
  group_by(Birth_Year) %>% 
  mutate(total = n(),
         rate = total/count_brth_yr*10000) %>% 
  select(Birth_Year, rate, Diag, cat) %>% 
  distinct() %>% 
  ungroup()
  
mod1 <- modeling_anal(anom, "q01")

if (knitr::is_html_output()){
p1 <- plot_line(data = dta,
                var = "rate")

p1 %>% 
  add_trace(data = mod1,
            x = ~Birth_Year,
            y = ~rate,
            name = ifelse(mod1[["model"]] %in% "nb",
                          "Negative Binomial GLM fit",
                          "Poisson GLM fit"),
            mode = "lines",
            line = list(color = "black"),
            marker = list(color = "black"),
         hovertemplate = ~paste(
                    "<b>", cat, "-", Birth_Year, "</b>",
                    "<br> Estimated Prevalence:",
                    scales::comma(rate,
                                  accuracy = 0.001),
                    "<extra></extra>" # removes the trace name from the hover text
                  )
         )
} else if(knitr::is_latex_output()){
  mod1 <- mod1 %>% 
  mutate(cat = ifelse(model %in% "nb",
         "Negative Binomial GLM fit",
         ifelse(model %in% "pois",
         "Poisson GLM fit", cat))) %>% 
    select(-c(model,total,count_brth_yr))
  
  new_dta <- rbind(dta, mod1)
  
  plot_line_pdf(data = new_dta,
                var = "rate")
}
```

```{r nt13, fig.cap=paste0(unique(anom %>% filter(tolower(Diag) %in% "q05") %>% select(cat) %>% distinct() %>% pull())," - Prevalence."),out.width=if(knitr::is_latex_output()){"65%"}}
dta <- subset(anom, tolower(Diag) %in% "q05") %>% 
  select(CaseID, Birth_Year, count_brth_yr, Diag, cat) %>% 
  distinct() %>% 
  group_by(Birth_Year) %>% 
  mutate(total = n(),
         rate = total/count_brth_yr*10000) %>% 
  select(Birth_Year, rate, Diag, cat) %>% 
  distinct() %>% 
  ungroup()
  
mod1 <- modeling_anal(anom, "q05")

if (knitr::is_html_output()){
p1 <- plot_line(data = dta,
                var = "rate")

p1 %>% 
  add_trace(data = mod1,
            x = ~Birth_Year,
            y = ~rate,
            name = ifelse(mod1[["model"]] %in% "nb",
                          "Negative Binomial GLM fit",
                          "Poisson GLM fit"),
            mode = "lines",
            line = list(color = "black"),
            marker = list(color = "black"),
         hovertemplate = ~paste(
                    "<b>", cat, "-", Birth_Year, "</b>",
                    "<br> Estimated Prevalence:",
                    scales::comma(rate,
                                  accuracy = 0.001),
                    "<extra></extra>" # removes the trace name from the hover text
                  )
         )
} else if(knitr::is_latex_output()){
  mod1 <- mod1 %>% 
  mutate(cat = ifelse(model %in% "nb",
         "Negative Binomial GLM fit",
         ifelse(model %in% "pois",
         "Poisson GLM fit", cat))) %>% 
    select(-c(model,total,count_brth_yr))
  
  new_dta <- rbind(dta, mod1)

plot_line_pdf(data = new_dta,
                var = "rate")

}
```

```{asis, echo=knitr::is_latex_output()}
\clearpage
```


## Selected central nervous system defects {#section33}

```{asis, echo=knitr::is_latex_output()}
\clearpage
```

## Selected sense organ defects {#section34}


```{asis, echo=knitr::is_latex_output()}
\clearpage
```

## Selected congenital heart defects {#section35}


```{asis, echo=knitr::is_latex_output()}
\clearpage
```

