---
fig.caption: yes
---

```{r setup, include=FALSE}
source("utilities/data.R")
source("utilities/plotLine.R")
source("utilities/CI.R")
source("utilities/modeling.R")
```

# Appendix {-}

The following table of selected sentinel anomalies indicate the trends in congenital anomaly prevalence in Nova Scotia from `r min(unique(anom_grp$BrthYear), na.rm = TRUE)` through `r max(unique(anom_grp$BrthYear), na.rm = TRUE)`. Sentinel anomalies are those which the International Clearinghouse of Birth Defects Surveillance and Research (ICBDSR), of which we are a member, watches worldwide with the rationale that they are quite easily identified hence more accurately reported. *See Appendix A.5 (p. 53) for other anomalies listed in the report.*

## Appendix 1 {#section-a1}

```{r trend-1}
##-------------------------
## neural tube defect (ntd)
##-------------------------

ntb <- modeling_trend(anom_grp, "q00")

## (sub categories)
ntb_sub <- lapply(
  c("q00",
    "q01",
    "q05"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

ntb_sub <- as.data.frame(do.call(rbind, ntb_sub))

##-----------------------
## Central nervous system
##-----------------------

cns <- modeling_trend(anom_grp, "q02")
cns_sub <- lapply(
  c("q02",
    "q03",
    "q042"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

cns_sub <- as.data.frame(do.call(rbind, cns_sub))

## ntd (sub categories)

##--------------------
## Sense organ defects
##--------------------

sod <- modeling_trend(anom_grp, "q110")

## (sub categories)
sod_sub <- lapply(
  c("q110",
    "q160",
    "q300"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

sod_sub <- as.data.frame(do.call(rbind, sod_sub))

##-------------------------
## Congenital heart defects
##-------------------------

chd <- modeling_trend(anom_grp, "q200")

## (sub categories)
chd_sub <- lapply(
  c("q200",
    "q2031",
    "q212",
    "q213",
    "q234"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

chd_sub <- as.data.frame(do.call(rbind, chd_sub))

##------------------
## Oro-facial clefts
##------------------

ofc <- modeling_trend(anom_grp, "q35")

## (sub categories)
ofc_sub <- lapply(
  c("q35",
    "q36",
    "q37"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

ofc_sub <- unique(as.data.frame(do.call(rbind, ofc_sub)))

##-------------------------
## Gastrointestinal defects
##-------------------------

gid <- modeling_trend(anom_grp, "q391")

## (sub categories)
gid_sub <- lapply(
  c("q390",
    "q41",
    "q420",
    "q431",
    "q442"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

gid_sub <- unique(as.data.frame(do.call(rbind, gid_sub)))

##------------------
## Genital anomalies
##------------------

ga <- modeling_trend(anom_grp, "q531")

## (sub categories)
ga_sub <- lapply(
  c("q531",
    "q54",
    "q56",
    "q640"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

ga_sub <- unique(as.data.frame(do.call(rbind, ga_sub)))

##----------------------
## Urinary tract defects
##----------------------

utd <- modeling_trend(anom_grp, "q601")

## (sub categories)
utd_sub <- lapply(
  c("q601",
    "q611",
    "q641",
    "q642"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

utd_sub <- unique(as.data.frame(do.call(rbind, utd_sub)))

##--------------
## Hip dysplasia
##--------------

hip <- modeling_trend(anom_grp, "q65")

##------------------------
## Limb deficiency defects
##------------------------

ldd <- modeling_trend(anom_grp, "q71")

##-----------------------
## Abdominal wall defects
##-----------------------

awd <- modeling_trend(anom_grp, "q792")

## (sub categories)
awd_sub <- lapply(
  c("q792",
    "q793"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

awd_sub <- unique(as.data.frame(do.call(rbind, awd_sub)))

##--------------------
## Chromosomal defects
##--------------------

cd <- modeling_trend(anom_grp, "q90")

## (sub categories)
cd_sub <- lapply(
  c("q90",
    "q914",
    "q910",
    "q96"),
  function(x) modeling_trend(data = anom_ind,
                             value = x))

cd_sub <- unique(as.data.frame(do.call(rbind, cd_sub)))

## putting them all together
dta <- rbind(
  ntb, ntb_sub,
  cns, cns_sub,
  sod, sod_sub,
  chd, chd_sub,
  ofc, ofc_sub,
  gid, gid_sub,
  ga, ga_sub,
  utd, utd_sub,
  hip,
  ldd,
  awd, awd_sub,
  cd, cd_sub
) %>% 
  mutate(pvalue = ifelse(!pvalue %in% "< 0.0001",
                         scales::comma(as.numeric(pvalue), accuracy = 0.001),
                         pvalue))

## printing table
DT::datatable(dta,
              rownames = FALSE,
              colnames = c("Coeficient",
                           "Selected anomaly",
                           "Number of periods",
                           "Trend direction",
                           "Statistic - \U03C7<sup>2</sup><sub>1</sub>",
                           "P-value"),
              escape = FALSE,
              # style = "bootstrap",
        selection = 'single',
        caption = " Linear trend analysis selected anomalies 1987–2021 inclusive",
        # width = "10px",
        extensions = "Buttons",
        options = list(
          # https://datatables.net/reference/option/dom
          dom = 'B<t>tp',
          extensions = "Buttons",
          # search = NULL,
          searchHighlight = TRUE,
          paging = TRUE,
          autoWidth = TRUE,
          pageLength = 10,
          ordering = TRUE,
          stateSave = TRUE,
          buttons = list('copy', 'print', list(
            extend = 'collection',
            buttons = list(
                list(extend = 'csv', filename = "trend_results")
                # list(extend = 'excel', filename = "trend_results"),
                # list(extend = 'pdf', filename = "trend_results")
                ),
            text = 'Download'
          )),
          columnDefs = list(list(visible = FALSE,
                                 targets = c(0,1)),
                            list(targets = c(0),
                                 searchable = TRUE))
          # language = list(zeroRecords = "")
        )) %>% 
  DT::formatRound(
          columns = c("statistic"),
          digits = 2,
          mark = ","
        )
```