---
fig.caption: yes
output: pdf_document
params:
  year1: 2013
  year2: 2022
  multi: 10000
---

```{r setup, include=FALSE}
source("utilities/data.R")
source("utilities/plotLine.R")
source("utilities/CI.R")
source("utilities/modeling.R")
```

# Appendix {.unnumbered}

The following table of selected sentinel anomalies indicate the trends in congenital anomaly prevalence in Nova Scotia from `r min(unique(anom$Birth_Year), na.rm = TRUE)` through `r max(unique(anom$Birth_Year), na.rm = TRUE)`. Sentinel anomalies are those which the International Clearinghouse of Birth Defects Surveillance and Research (ICBDSR), of which we are a member, watches worldwide with the rationale that they are quite easily identified hence more accurately reported. *See Section* \@ref(section21) *for all the anomalies listed in the report.*

## Appendix 1 - Linear trend analysis and p-values, `r paste0(params$year1,"-",params$year2)` {#section-a1}

```{r trend1, fig.cap=""}
## neural tube defect (ntd) ----

dta <- anom %>%
  filter(!SrceIDs %in% 5, #remove MSI
         between(Birth_Year, params$year1, params$year2)) %>%
  select(CaseID, Birth_Year, Diag, cat, count_brth_yr) %>% distinct()

ntb <- modeling_trend(dta, "gr101")

## (sub categories)
ntb_sub <- lapply(
  c("q00",
    "q01",
    "q05"),
  function(x) modeling_trend(dta,
                             value = x))

ntb_sub <- as.data.frame(do.call(rbind, ntb_sub))

## Central nervous system ----

cns <- modeling_trend(dta, "gr102")
cns_sub <- lapply(
  c("q02",
    "q03",
    "q041"),
  function(x) modeling_trend(dta,
                             value = x))

cns_sub <- as.data.frame(do.call(rbind, cns_sub))

## Sense organ defects ----

sod <- modeling_trend(dta, "gr103")

## (sub categories)
sod_sub <- lapply(
  c("q110",
    "q160",
    "q300"),
  function(x) modeling_trend(dta,
                             value = x))

sod_sub <- as.data.frame(do.call(rbind, sod_sub))

## Congenital heart defects ----

chd <- modeling_trend(dta, "gr104")

## (sub categories)
chd_sub <- lapply(
  c("q200",
    "q201",
    "q212",
    "q213",
    "q234",
    "q251"),
  function(x) modeling_trend(dta,
                             value = x))

chd_sub <- as.data.frame(do.call(rbind, chd_sub))

## Oro-facial clefts ----

ofc <- modeling_trend(dta, "gr105")

## (sub categories)
ofc_sub <- lapply(
  c("q35",
    "q36",
    "q37"),
  function(x) modeling_trend(dta,
                             value = x))

ofc_sub <- unique(as.data.frame(do.call(rbind, ofc_sub)))

## Gastrointestinal defects ----

gid <- modeling_trend(dta, "gr106")

## (sub categories)
gid_sub <- lapply(
  c("q390",
    "q41",
    "q420",
    "q431",
    "q442"),
  function(x) modeling_trend(dta,
                             value = x))

gid_sub <- unique(as.data.frame(do.call(rbind, gid_sub)))

## Genital anomalies ----

ga <- modeling_trend(dta, "gr107")

## (sub categories)
ga_sub <- lapply(
  c("q531",
    "q54",
    "q56",
    "q640"),
  function(x) modeling_trend(dta,
                             value = x))

ga_sub <- unique(as.data.frame(do.call(rbind, ga_sub)))

## Urinary tract defects ----

utd <- modeling_trend(dta, "gr108")

## (sub categories)
utd_sub <- lapply(
  c("q600",
    "q611",
    "q641",
    "q642"),
  function(x) modeling_trend(dta,
                             value = x))

utd_sub <- unique(as.data.frame(do.call(rbind, utd_sub)))

## Hip dysplasia ----

hip <- modeling_trend(dta, "gr109")

## Limb deficiency defects ----

ldd <- modeling_trend(dta, "gr110")

## Abdominal wall defects ----

awd <- modeling_trend(dta, "gr111")

## (sub categories)
awd_sub <- lapply(
  c("q792",
    "q793"),
  function(x) modeling_trend(dta,
                             value = x))

awd_sub <- unique(as.data.frame(do.call(rbind, awd_sub)))

## Chromosomal defects ----

cd <- modeling_trend(dta, "gr112")

## (sub categories)
cd_sub <- lapply(
  c("q90",
    "q914",
    "q910",
    "q96"),
  function(x) modeling_trend(dta,
                             value = x))

cd_sub <- unique(as.data.frame(do.call(rbind, cd_sub)))

## putting them all together
dta <- rbind(
  ntb, ntb_sub,
  cns, cns_sub,
  sod, sod_sub,
  chd, chd_sub,
  ofc, ofc_sub,
  gid, gid_sub,
  ga, ga_sub,
  utd, utd_sub,
  hip,
  ldd,
  awd, awd_sub,
  cd, cd_sub
) %>% 
  mutate(pvalue = ifelse(pvalue < 0.000199 ,
                         "< 0.0001",
                         scales::comma(as.numeric(pvalue), accuracy = 0.001)),
         statistic = as.numeric(statistic)) %>% 
  select(-c(coef, std, rate))

if(knitr::is_html_output()){
## printing table
DT::datatable(dta,
              rownames = FALSE,
              colnames = c(#"Coeficient",
                           "Selected anomaly",
                           "Number of cases",
                           "Trend direction",
                           "Statistic <br> (\U03C7<sup>2</sup><sub>1</sub>)",
                           "P-value"),
              escape = FALSE,
              # style = "bootstrap",
        selection = 'single',
        # caption = "Linear trend analysis selected anomalies 1987-2021 inclusive",
        # width = "10px",
        extensions = "Buttons",
        options = list(
          # https://datatables.net/reference/option/dom
          dom = 'B<t>tp',
          extensions = "Buttons",
          # search = NULL,
          searchHighlight = TRUE,
          paging = TRUE,
          autoWidth = TRUE,
          pageLength = 10,
          ordering = TRUE,
          stateSave = TRUE,
          buttons = list('copy', 'print', list(
            extend = 'collection',
            buttons = list(
                list(extend = 'csv', filename = "trend_results")
                # list(extend = 'excel', filename = "trend_results"),
                # list(extend = 'pdf', filename = "trend_results")
                ),
            text = 'Download'
          )),
          columnDefs = list(#list(visible = FALSE,
                                # targets = c(0)),
                            list(targets = c(1),
                                 searchable = TRUE))
          # language = list(zeroRecords = "")
        )) %>% 
  DT::formatRound(
          columns = c("statistic"),
          digits = 2,
          mark = ","
        )
} else if(knitr::is_latex_output()){
  dta %>%
    mutate(anom = gsub("<br>","\n", anom),
           anom = gsub("<span style='font-size:12px'>",
                       " \\\\tiny{", anom),
           anom = gsub("</span>", "}", anom),
           anom = linebreak(anom, align = "l"),
           n = gsub("<br>","\n", n),
           n = gsub("<span style='font-size:12px'>",
                       " \\\\tiny{", n),
           n = gsub("</span>", "}", n),
           n = linebreak(n, align = "c"),
           trend = ifelse(grepl("increasing", tolower(trend)),
                          "Increasing $\\left(\\uparrow \\right)$",
                   ifelse(grepl("decreasing", tolower(trend)),
                          "Decreasing $\\left(\\downarrow \\right)$","No significant \n change $\\left(\\rightarrow \\right)$")),
           trend = linebreak(trend, align = "c")
           ) %>%
    kableExtra::kbl(
      format = "latex",
      escape = FALSE,
      booktabs = TRUE,
      longtable = TRUE,
      # table.attr = "style='width:65%;'",
      align = c("l","c","c","c","c"),
      col.names = c("Selected anomaly",
                    linebreak("Number of\n cases", align = "c"),
                    linebreak("Trend \n direction", align = "c"),
                    linebreak("Statistic \n $\\left(\\chi^{2}_{1}\\right)$", align = "c"),
                    "P-value")
      ) %>% 
    kable_styling(latex_options = c("repeat_header", "stripped"), font_size = 8) %>% 
    column_spec(1, width = "4cm")
    
    
}
```

```{asis, echo=knitr::is_latex_output()}
\clearpage
```

## Appendix 2 - Total birth, number of cases, anomalies and anomalies per case, `r paste0(params$year1,"-",params$year2)` {#section-a2}

```{r trend2, fig.cap=""}

## Number of cases ----
dta <- anom %>% 
  filter(!grepl("gr|q999",tolower(Diag)),
         !SrceIDs %in% 5, #remove MSI
         between(Birth_Year, params$year1, params$year2)) %>%
  select(CaseID, Birth_Year, count_brth_yr, Diag) %>% 
  select(CaseID, Birth_Year, count_brth_yr) %>%
  distinct() %>% 
  group_by(Birth_Year) %>% 
  mutate(cases = n()) %>% 
  select(Birth_Year, cases, count_brth_yr) %>% 
  distinct() %>% 
  ungroup()

dta1 <- anom %>% 
  filter(!SrceIDs %in% 5, #remove MSI
         !grepl("gr|q999",tolower(Diag))) %>%
  select(CaseID, Birth_Year, Diag) %>% 
  distinct() %>% 
  group_by(Birth_Year) %>% 
  mutate(anom = n()) %>% 
  select(Birth_Year, anom) %>% 
  distinct() %>% 
  ungroup()

dta <- merge(dta,
             dta1,
             by = "Birth_Year")

dta <- dta %>% 
  mutate(
    case_rate = params$multi*(cases/count_brth_yr),
    anom_rate = params$multi*(anom/count_brth_yr),
    avg = anom/cases) %>% 
  select(Birth_Year, count_brth_yr, cases, case_rate, anom, anom_rate, avg)

if(knitr::is_html_output()){
## printing table
DT::datatable(dta,
              rownames = FALSE,
              colnames = c("Year",
                           "Total births",
                           "Number of cases",
                           "Case rate <br> per 10,000 total births",
                           "Number of anomalies",
                           "Anomaly rate <br> per 10,000 total births",
                           "Anomaly rate per case"),
              escape = FALSE,
              # style = "bootstrap",
        selection = 'single',
        # caption = "Linear trend analysis selected anomalies 1987-2021 inclusive",
        # width = "10px",
        extensions = "Buttons",
        options = list(
          # https://datatables.net/reference/option/dom
          dom = 'B<t>tp',
          extensions = "Buttons",
          # search = NULL,
          searchHighlight = TRUE,
          paging = TRUE,
          autoWidth = TRUE,
          pageLength = 10,
          ordering = TRUE,
          stateSave = TRUE,
          buttons = list('copy', 'print', list(
            extend = 'collection',
            buttons = list(
                list(extend = 'csv', filename = "anomaly_summary")
                # list(extend = 'excel', filename = "trend_results"),
                # list(extend = 'pdf', filename = "trend_results")
                ),
            text = 'Download'
          )),
          columnDefs = list(#list(visible = FALSE,
                                # targets = c(0)),
                            list(targets = c(1),
                                 searchable = TRUE))
          # language = list(zeroRecords = "")
        )) %>% 
  DT::formatRound(
          columns = c("case_rate","anom_rate","avg"),
          digits = 2,
          mark = ","
        )
} else if(knitr::is_latex_output()){
  dta %>%
    mutate(cases = scales::comma(cases, accuracy = 1),
           anom = scales::comma(anom, accuracy = 1),
           case_rate = scales::comma(case_rate, accuracy = 0.01),
           anom_rate = scales::comma(anom_rate, accuracy = 0.01),
           avg = scales::comma(avg, accuracy = 0.01)) %>%
    kableExtra::kbl(
      format = "latex",
      escape = FALSE,
      booktabs = TRUE,
      longtable = TRUE,
      # table.attr = "style='width:65%;'",
      align = c("c","c","c","c","c","c","c"),
      col.names = c("Year",
                    linebreak("Total \n births", align = "c"),
                    linebreak("Number of \n cases", align = "c"),
                    linebreak("Case rate per \n 10,000 total births", align = "c"),
                    linebreak("Number of \n anomalies", align = "c"),
                    linebreak("Anomaly rate per \n 10,000 total births", align = "c"),
                    linebreak("Anomaly rate \n per case", align = "c"))
      ) %>% 
    kable_styling(latex_options = c("repeat_header", "striped"), font_size = 8)
}
```

## Appendix 3 - Prevalence rate per 10,000 total births, aggregated prevalence rate, `r paste0(params$year1,"-",params$year2)` {#section-a23}

```{r trend3, fig.cap=""}

ord <- c("Neural tube defects", "(Q00) - Anencephaly and similar malformations", "(Q01) - Encephalocele", "(Q05) - Spina bifida", "Selected central nervous system defects", "(Q02) - Microcephaly", "(Q03) - Congenital hydrocephalus", "(Q04.1, Q04.2) - Arhinencephaly / Holoprosencephaly", "Selected sense organ defects", "(Q11.0-Q11.2) - Anophtalmos / Microphtalmos", "(Q16.0, Q17.2) - Anotia / Microtia", "(Q30.0) - Choanal atresia", "Selected congenital heart defects", "(Q20.0) - Commom truncus", "(Q20.1, Q20.3, Q20.5) - Transposition of great vessels", "(Q21.2) - Atrioventricular septal defect", "(Q21.3) - Tetralogy of Fallot", "(Q23.4) - Hypoplastic left heart syndrome", "(Q25.1) - Coarctation of aorta", "Oro-facial clefts", "(Q35) - Cleft palate only", "(Q36) - Cleft lip only", "(Q37) - Cleft palate with cleft lip", "Selected gastrointestinal defects", "(Q39.0-Q39.4) - Oesophageal atresia / stenosis, tracheoesophageal fistula", "(Q41) - Small intestine absence / atresia / stenosis", "(Q42.0-Q42.3) - Ano-rectal absence / atresia / stenosis", "(Q43.1) - Hirschsprung disease", "(Q44.2) - Atresia of bile ducts", "Selected genital anomalies", "(Q53.1, Q53.2, Q53.9) - Cryptorchidism / undescended testicles", "(Q54, excluding Q54.4) - Hypospadias", "(Q56) - Indeterminate sex", "(Q64.0) - Epispadias", "Selected urinary tract defects", "(Q60.0-Q60.2) - Renal agenesis", "(Q61.1-Q61.5, Q61.8, Q61.9) - Cystic kidney", "(Q64.1) - Bladder and cloacal exstrophy", "(Q64.2, Q64.3) - Lower urinary tract obstruction", "Hip dysplasia", "(Q65) - Hip dysplasia", "Limb deficiency defects", "(Q79.0) - Diaphragmatic hernia", "(Q71.4-Q71.9, Q72.4-Q72.9, Q73.8, excluding Q71.6, Q71.7, Q72.7) - Limb deficiency defects", "Selected abdominal wall defects", "(Q79.2) - Omphalocele / Exomphalos", "(Q79.3) - Gastroschisis", "Selected chromosomal defects", "(Q90) - Down Syndrome", "(Q91.4-Q91.7) - Trisomy 13 - Patau", "(Q91.0-Q91.3) - Trisomy 18 - Edwards", "(Q96) - Turner syndrome")

## Number of cases ----
dta <- anom %>% 
  filter(!SrceIDs %in% 5, #remove MSI
         between(Birth_Year, params$year1, params$year2),
         !grepl("q999",tolower(Diag))) %>%
  select(CaseID, Birth_Year, count_brth_yr, Diag, cat) %>% 
  distinct() %>% 
  group_by(Birth_Year, Diag) %>% 
  mutate(cases = n(),
         rate = params$mult*cases/count_brth_yr,
         cat = factor(cat, levels = ord)) %>% 
  ungroup() %>% 
  arrange(cat) %>% 
  select(cat, rate) %>% 
  distinct()

dta <- split(dta$rate, dta$cat)

dta1 <- anom %>% 
  filter(!SrceIDs %in% 5, #remove MSI
         between(Birth_Year, params$year1, params$year2),
         !grepl("q999",tolower(Diag))) %>%
  select(CaseID, Birth_Year, count_brth_yr, Diag, cat) %>% 
  distinct() %>% 
  group_by(Diag) %>% 
  mutate(cases = n(),
         Birth_Year = paste0(params$year1, "-", params$year2),
         count_brth_yr = sum(count_brth_yr, na.rm = TRUE),
         rate = params$multi*cases/count_brth_yr,
         cat = factor(cat, levels = ord),
         plot = "")%>% 
  ungroup() %>% 
  arrange(cat) %>% 
  select(cat, cases, rate, plot) %>% 
  distinct()

  dta1 %>%
    mutate(rate = scales::comma(rate, accuracy = 0.001),
           cases = scales::comma(cases, accuracy = 1)) %>%
    kableExtra::kbl(
      format = "latex",
      escape = FALSE,
      booktabs = TRUE,
      longtable = TRUE,
      # table.attr = "style='width:65%;'",
      align = c("l","c","c","c","c","c","c"),
      col.names = c(
        linebreak("Selected anomaly", align = "l"),
        linebreak("Number of \n cases", align = "c"),
        linebreak("Aggregated prevalence rate per \n 10,000 total births \n (2012-2022)", align = "c"),
        linebreak("Prevalence rate per \n 10,000 total births \n (2012-2022)", align = "c")
        )
      ) %>%
    kable_styling(latex_options = c("repeat_header", "striped"), font_size = 8) %>% 
    column_spec(4, image = spec_plot(dta, same_lim = FALSE, type = "l"),
                latex_valign = "m") %>% 
    column_spec(1, width = "4cm", latex_valign = "m")
```